---
title: "Plant - Red List Dashboard"
author: By Steven Bachman 
output: 
  flexdashboard::flex_dashboard:
  #orientation: rows
    theme: cosmo
    source_code: embed
vertical_layout: fill
runtime: shiny
---
  
  
```{r setup, include=FALSE}
library(flexdashboard)
library(magrittr)
library(readr)
library(ggplot2)
library(plotly)
library(rgdal)
library(DT)
library(leaflet)

# load in the Red List data
all_plants_sp <-
read_csv("all.plants.RL_full.csv", guess_max = 50000)

# change the levels to remove old categories
all_plants_sp$category[all_plants_sp$category == "LR/nt"] = "NT"
all_plants_sp$category[all_plants_sp$category == "LR/cd"] = "LC"
all_plants_sp$category[all_plants_sp$category == "LR/lc"] = "LC"

# set order of levels
all_plants_sp$category = factor(all_plants_sp$category,
levels = c("EX", "EW", "CR", "EN", "VU", "NT", "LC", "DD"))

# set groups
All_plants = all_plants_sp

Land_plants = subset(
all_plants_sp,
all_plants_sp$phylum == "TRACHEOPHYTA" |
all_plants_sp$phylum == "ANTHOCERATOPHYTA" |
all_plants_sp$phylum == "ANTHOCERATOPHYTA" |
all_plants_sp$phylum == "MARCHANTIOPHYTA"
)

Vascular_plants =  subset(all_plants_sp, all_plants_sp$phylum == "TRACHEOPHYTA")

Seed_plants = subset(
all_plants_sp,
all_plants_sp$class == "CYCADOPSIDA" |
all_plants_sp$class == "GINKGOOPSIDA" |
all_plants_sp$class == "GNETOPSIDA" |
all_plants_sp$class == "PINOPSIDA" |
all_plants_sp$class == "LILIOPSIDA" |
all_plants_sp$class == "MAGNOLIOPSIDA"
)

Flowering_plants = subset(
all_plants_sp,
all_plants_sp$class == "LILIOPSIDA" |
all_plants_sp$class == "MAGNOLIOPSIDA"
)

Gymnosperms = subset(
all_plants_sp,
all_plants_sp$class == "CYCADOPSIDA" |
all_plants_sp$class == "GINKGOOPSIDA" |
all_plants_sp$class == "GNETOPSIDA" |
all_plants_sp$class == "PINOPSIDA"
)

Green_Algae = subset(
all_plants_sp,
all_plants_sp$phylum == "CHAROPHYTA" |
all_plants_sp$phylum == "CHLOROPHYTA"
)

Mosses = subset(
all_plants_sp,
all_plants_sp$phylum == "BRYOPHYTA" |
all_plants_sp$phylum == "MARCHANTIOPHYTA" |
all_plants_sp$phylum == "ANTHOCERATOPHYTA"
)

Red_Algae =  subset(all_plants_sp, all_plants_sp$phylum == "RHODOPHYTA")

Ferns_and_Allies = subset(
all_plants_sp,
all_plants_sp$class == "POLYPODIOPSIDA" |
all_plants_sp$class == "LYCOPODIOPSIDA"
)

# number of described plants using different gorupings - from Nic Lughadha 2016 Counting Counts
All_plants_tot = 450000 # work this out
Land_plants_tot = 403911
Vascular_plants_tot = 383671
Seed_plants_tot = 370492
Flowering_plants_tot = 369434
Gymnosperms_tot = 1200 # work this out
Green_Algae_tot = 500 # work this out
Mosses_tot = 500 # work this out
Red_Algae_tot = 500 # work this out
Ferns_and_Allies_tot = 500 # work this out
```

Global status {data-orientation=columns}
===================================== 
  Inputs{.sidebar}
-------------------------------------
  
  Red List Version 2018-1

- [IUCN Red List](http://www.iucnredlist.org/)




```{r}
# Add options to the select input widget

selectInput(
  "Group",
  label = ("Select plant group"),
  choices = list(
    "All plants" = "All_plants" ,
    "Land plants" = "Land_plants" ,
    "Vascular plants" = "Vascular_plants",
    "Seed plants" = "Seed_plants",
    "Flowering plants" = "Flowering_plants",
    "Gymnosperms" = "Gymnosperms",
    "Ferns and Allies" = "Ferns_and_Allies",
    "Mosses" = "Mosses",
    "Green Algae" = "Green_Algae",
    "Red Algae" = "Red_Algae"
  ),
  selected = "All_plants"
)

# Configure slider widget

sliderInput(
  "sliderYear",
  label = "Select year range",
  min = 1998,
  max = 2018,
  value = c(1998, 2018),
  sep = ""
)
```





Column {data-width=700}
-------------------------------------
### Count of species per Red List category
  
```{r}
#  

renderPlot({

# get the selection group    
myinput = get(input$Group)

# build the plot with selection group and combine with year slider selection
ggplot(myinput[which(
  myinput$published_year >= input$sliderYear[1] &
  myinput$published_year <= input$sliderYear[2]
  ),],
  aes(category)) +
  geom_bar(aes(fill = category)) +
  theme_minimal() +
  scale_fill_manual(
  values = c(
  "#080100",
  "#4A235A",
  "#d5392a",
  "#D35400",
  "#F1C40F",
  "#A9DFBF",
  "#1E8449",
  "#AAAAAA"
  ),
  labels = c("EX", "EW", "CR", "EN", "VU", "NT", "LC", "DD"),
  drop = FALSE
  )
})
```

Column {data-width=400}
-------------------------------------
  Summary statistics

### Species assessments

```{r}
# get total number of species assessments and render as a Value Box
renderValueBox({
  myinput = get(input$Group)
  get_group = myinput[which(
  myinput$published_year >= input$sliderYear[1] &
  myinput$published_year <= input$sliderYear[2]
  ),]
  group_rows = nrow(get_group)
  valueBox(prettyNum(group_rows, big.mark = ","), color = "#acacac")
})
```

### % of group assessed

```{r}
# get total number of species assessments and divide by total # in group to get %, then render as a Value Box
renderValueBox({
  myinput = get(input$Group)
  mytotal = paste0(input$Group, "_tot")
  mytotal = get(mytotal)
  get_group = myinput
  group_rows = round((nrow(get_group) / mytotal) * 100, digits = 1)
  valueBox(group_rows, icon = "percentage", color = "#acacac")
})
```


### % threatened (CR + EN + VU)

```{r}
# get total number of threatened assessments as % of total and render as a Value Box
renderValueBox({
  myinput = get(input$Group)
  get_group_T = myinput[which(
  myinput$published_year >= input$sliderYear[1] &
  myinput$published_year <= input$sliderYear[2]
  ),]
  group_rows_T = nrow(get_group_T)
  threatened = subset(get_group_T, category == "CR" |
  category == "EN" | category == "VU")
  nrow_threatened = nrow(threatened)
  percent_T = round((nrow_threatened / group_rows_T) * 100, digits = 1)
  myicon = ifelse(percent_T > 20, "fa-arrow-up", "arrows-alt-h")
  mycolour = ifelse(percent_T > 20, "#d5392a", "#acacac")
  val = percent_T
  valueBox(val,
  color =  mycolour)
})
```


### % Data Deficient

```{r}
# get total number of data deficient assessments as % of total and render as a Value Box
renderValueBox({
  myinput = get(input$Group)
  get_group_DD = myinput[which(
  myinput$published_year >= input$sliderYear[1] &
  myinput$published_year <= input$sliderYear[2]
  ), ]
  get_group_onlyDD = get_group_DD[which(get_group_DD$category == "DD"), ]
  group_rows_DD = nrow(get_group_DD)
  group_rows_onlyDD = nrow(get_group_onlyDD)
  proportion = round((group_rows_onlyDD / group_rows_DD) * 100, digits = 1)
  myicon = ifelse(proportion > 40, "fa-arrow-up", "arrows-alt-h")
  mycolour = ifelse(proportion > 40, "#d5392a", "#acacac")
  val = proportion
  valueBox(val,
  #icon = myicon,
  color =  mycolour)
})
```

### Increase in number of assessments since `r max(all_plants_sp$published_year)-1 `

```{r}
# get total number of assessments since last year
renderValueBox({
  get_group = get(input$Group)
  group_rows = nrow(subset(get_group, subset = published_year == max(published_year)))
  valueBox(prettyNum(group_rows, big.mark = ","),
  icon = "fa-arrow-up",
  color = "#acacac")
})
```

### Number of assessments > 10 years old

```{r}
# get total number of assessments more than 10 years since maximum year
renderValueBox({
  get_group = get(input$Group)
  group_rows = nrow(subset(get_group, subset = published_year < max(published_year) -
  10))
  valueBox(prettyNum(group_rows, big.mark = ","), color = "#acacac")
})
```

Trends {data-orientation=rows}
=====================================     
  
  Inputs{.sidebar}
-------------------------------------
  
  Red List Version 2018-1

- [IUCN Red List](http://www.iucnredlist.org/)




**Left tab** shows count of assessments per year

**Middle tab** shows breakdown of assessments categorised as 'new' (assessed within 5 years of being described) and 'old' (assessed > 5 years after description)

**Right tab** shows trends in use of different assessment languages






Row {data-height=600 .tabset .tabset-fade}
-------------------------------------
### Assessment count per year
  
```{r, }
# Count of Red List Assessments for plant species over time 
# filter out <2003 (old assessments, old criteria) and >2017 (2018 not finished yet)
all_plants_sp_pubyear = subset(all_plants_sp, published_year > 2002 &
                                 published_year < 2018)
                                 
# plot as a bar chart
ggplotly(
  ggplot(all_plants_sp_pubyear, aes(as.factor(published_year))) +
    geom_bar(fill = "#FF6666") +
    theme_minimal() +
    labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "")
)
```

### Species description vs Red List assessment date

```{r, }
# load in the data on 
diffplot <- read_csv("diffplot.csv", guess_max = 50000)

# adding in a single record to get a point for 2005 - must be a better way to do this...
extralev = data.frame(published_year = "2005", age = "Unknown")
diffplot = rbind(diffplot, extralev)

# subset the data on number of new species described
#tax.nov.plot = subset(tax.nov.year, tax.nov.year$Year >="2003" & tax.nov.year$Year < "2018")

# now summarise the ratio of old and new for each red list assessment year
fig_2 = ggplot() +
  geom_bar(
    data = diffplot,
    aes(x = as.factor(published_year), fill = as.factor(age)),
    position = position_dodge(preserve = 'single')
  ) +
  
  labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "") +
  theme_minimal() +
  scale_fill_manual(values = c(
    "New" = "#f37735",
    "Old" = "#ffc425",
    "Unknown" = "light grey"
  )) +
  scale_x_discrete(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 100))
ggplotly(fig_2)
```


### Assessment language
```{r, }
# import the Red List download
assessments_plants_2018_1 <-
read_csv("assessments_plants_2018_1.csv", guess_max = 50000)

# assign the NAs as English
assessments_plants_2018_1$language[is.na(assessments_plants_2018_1$language)] <-
"English"

# query out pre 2003, but leave in 2018 as this includes french 
assessments_plants_2018_1 = subset(assessments_plants_2018_1, yearPublished >
2002)

# plot the graph
cbPalette <-
c(
"#999999",
"#E69F00",
"#56B4E9",
"#009E73",
"#F0E442",
"#0072B2",
"#D55E00",
"#CC79A7"
)
rl_language = ggplot(data = assessments_plants_2018_1, aes(x = factor(yearPublished), fill = language))
rl_language = rl_language + geom_bar() +
theme_minimal() +
scale_fill_manual(values = cbPalette) +
labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "")

ggplotly(rl_language)

```


Specialist groups {data-orientation=rows}
===================================== 

Inputs{.sidebar}
-------------------------------------
  Red List Version 2018-1

- [IUCN Red List](http://www.iucnredlist.org/)


```{r, include=FALSE}
all_sg_rls = readOGR(dsn = "shapefiles", layer = "all_SG_RLA")
```

Row {data-height=600}
-------------------------------------
```{r}
# import the 
shared_SGs = all_sg_rls

renderLeaflet({
  
  leaflet(data = shared_SGs) %>%
    addTiles()  %>%
    setView(lat=10, lng=0, zoom=2) %>%
    addPolygons(fillColor = ~Group,
                popup = paste(shared_SGs$Group, "<br>", 
                              "<b><a href=",shared_SGs$Website,">Web link</a></b>", "<br>", 
                              "<b>Chair: </b>", shared_SGs$Chair,"<br>",
                              "<b>Chair email: </b>", shared_SGs$Email,"<br>",
                              "<b>RLA: </b>", shared_SGs$RLA,"<br>",
                              "<b>RLA email: </b>", shared_SGs$RLA_email,"<br>"),
                #label = ~Group,
                stroke = TRUE, 
                color = "black",
                weight = 2,
                fillOpacity = 0.2,
                highlight = highlightOptions(weight = 5,
                                             color = "red",
                                             fillOpacity = 0.7,
                                             bringToFront = TRUE))
  
})

```

Row {data-height=600}
-------------------------------------
  
### Taxonomic and thematic specialist group information
  
```{r}
# import the specialist group data
PlantSGsRLAs <- read.csv("PlantSGsRLAs.csv", encoding = "UTF-16")

# finish page with table view - DT seems to work well
datatable(PlantSGsRLAs, escape = FALSE)
```
