---
title: "Plant - Red List Dashboard"
author: By Steven Bachman 
output: 
  flexdashboard::flex_dashboard:
  #orientation: rows
  vertical_layout: fill
theme: yeti
source_code: embed
runtime: shiny
---
  
  
```{r setup, include=FALSE}
#deployApp()
library(flexdashboard)
library(magrittr)
library(readr)
library(ggplot2)
library(plotly)

# load data in 'global' chunk so it can be shared by all users of the dashboard
#all_plants_sp <- read_csv("all.plants.RL_full.csv")
all.plants.RL_full <- read_csv("all.plants.RL_full.csv", guess_max = 50000)

all_plants_sp = all.plants.RL_full


# change the levels to remove old categories
all_plants_sp$category[all_plants_sp$category=="LR/nt"] = "NT"
all_plants_sp$category[all_plants_sp$category=="LR/cd"] = "LC"
all_plants_sp$category[all_plants_sp$category=="LR/lc"] = "LC"

# set order of levels
all_plants_sp$category = factor(all_plants_sp$category, levels = c("EX", "EW", "CR", "EN", "VU", "NT", "LC", "DD"))

# set groups
All_plants = all_plants_sp

Land_plants = subset(all_plants_sp, all_plants_sp$phylum == "TRACHEOPHYTA" | 
                       all_plants_sp$phylum == "ANTHOCERATOPHYTA" | 
                       all_plants_sp$phylum == "ANTHOCERATOPHYTA" | 
                       all_plants_sp$phylum == "MARCHANTIOPHYTA")

Vascular_plants =  subset(all_plants_sp, all_plants_sp$phylum == "TRACHEOPHYTA" )

Seed_plants = subset(all_plants_sp, all_plants_sp$class == "CYCADOPSIDA" |
                       all_plants_sp$class == "GINKGOOPSIDA" |
                       all_plants_sp$class == "GNETOPSIDA" |
                       all_plants_sp$class == "PINOPSIDA" |
                       all_plants_sp$class == "LILIOPSIDA" |
                       all_plants_sp$class == "MAGNOLIOPSIDA")

Flowering_plants = subset(all_plants_sp, all_plants_sp$class == "LILIOPSIDA" |
                            all_plants_sp$class == "MAGNOLIOPSIDA" )

Gymnosperms = subset(all_plants_sp, all_plants_sp$class == "CYCADOPSIDA" |
                       all_plants_sp$class == "GINKGOOPSIDA" |
                       all_plants_sp$class == "GNETOPSIDA" |
                       all_plants_sp$class == "PINOPSIDA" )

Green_Algae = subset(all_plants_sp, all_plants_sp$phylum == "CHAROPHYTA" |
                       all_plants_sp$phylum == "CHLOROPHYTA")

Mosses = subset(all_plants_sp, all_plants_sp$phylum == "BRYOPHYTA" |
                  all_plants_sp$phylum == "MARCHANTIOPHYTA" |
                  all_plants_sp$phylum == "ANTHOCERATOPHYTA")

Red_Algae =  subset(all_plants_sp, all_plants_sp$phylum == "RHODOPHYTA" )

Ferns_and_Allies = subset(all_plants_sp, all_plants_sp$class == "POLYPODIOPSIDA" |
                            all_plants_sp$class == "LYCOPODIOPSIDA")

# number of described plants using different gorupings - from Nic Lughadha 2016 Counting Counts
All_plants_tot = 450000 # work this out
Land_plants_tot = 403911
Vascular_plants_tot = 383671
Seed_plants_tot = 370492
Flowering_plants_tot = 369434
Gymnosperms_tot = 1200 # work this out
Green_Algae_tot = 500 # work this out
Mosses_tot = 500 # work this out
Red_Algae_tot = 500 # work this out
Ferns_and_Allies_tot = 500 # work this out
```

Global status {data-orientation=columns}
===================================== 
  Inputs{.sidebar}
-------------------------------------
  
  Red List Version 2018-1

- [IUCN Red List](http://www.iucnredlist.org/)




```{r}
#multi select

selectInput("Group", label = ("Select plant group"), choices = list("All plants" = "All_plants" , "Land plants" = "Land_plants" , "Vascular plants" = "Vascular_plants", "Seed plants" = "Seed_plants", "Flowering plants" = "Flowering_plants", "Gymnosperms" = "Gymnosperms","Ferns and Allies" = "Ferns_and_Allies", "Mosses" = "Mosses","Green Algae" = "Green_Algae","Red Algae" = "Red_Algae"), 
            selected = "All_plants")

sliderInput("sliderYear", label = "Select year range", min = 1998, 
            max = 2018, value = c(1998, 2018), sep = "")
```



Bar chart shows number of species in each Red List category for selected group.

Column {data-width=700}
-------------------------------------
### Count of species per Red List category
  
```{r}

renderPlot({
  
  myinput = get(input$Group)
  
  ggplot(myinput[which(myinput$published_year >= input$sliderYear[1] & 
                         myinput$published_year <= input$sliderYear[2]),], 
         aes(category)) + 
    geom_bar(aes(fill = category)) +
    theme_minimal() +
    scale_fill_manual(
      values = c("#080100", "#4A235A", "#d5392a", "#D35400", "#F1C40F", "#A9DFBF", "#1E8449", "#AAAAAA"),
      labels = c("EX", "EW", "CR", "EN", "VU", "NT", "LC", "DD"),
      drop = FALSE)
})
```

Column {data-width=400}
-------------------------------------
  Summary statistics

### Species assessments

```{r}
# get total 
renderValueBox({
  myinput = get(input$Group)
  get_group = myinput[which(myinput$published_year >= input$sliderYear[1] & 
                              myinput$published_year <= input$sliderYear[2]),]
  group_rows = nrow(get_group)
  valueBox(prettyNum(group_rows, big.mark = ","), color = "#acacac")
})
```

### % of group assessed

```{r}
renderValueBox({
  myinput = get(input$Group)
  mytotal = paste0(input$Group,"_tot")
  mytotal = get(mytotal)
  get_group = myinput
  
  #percent_T = round((nrow_threatened/group_rows_T)*100, digits = 1)
  group_rows = round((nrow(get_group)/mytotal)*100, digits = 1)
  
  valueBox(group_rows, icon = "percentage", color = "#acacac")
})
```


### % threatened (CR + EN + VU)

```{r}
renderValueBox({
  myinput = get(input$Group)
  get_group_T = myinput[which(myinput$published_year >= input$sliderYear[1] & 
                                myinput$published_year <= input$sliderYear[2]),]
  group_rows_T = nrow(get_group_T)
  threatened = subset(get_group_T, category == "CR"|category == "EN"|category =="VU")
  nrow_threatened = nrow(threatened)
  #threatened + (threatened/non-DD) * DD
  percent_T = round((nrow_threatened/group_rows_T)*100, digits = 1)
  
  myicon = ifelse(percent_T > 20, "fa-arrow-up","arrows-alt-h")
  mycolour = ifelse(percent_T > 20, "#d5392a", "#acacac")
  val = percent_T
  
  valueBox(val, 
           #icon = myicon, 
           color =  mycolour)
})
```


### % Data Deficient

```{r}
renderValueBox({
  myinput = get(input$Group)
  get_group_DD = myinput[which(myinput$published_year >= input$sliderYear[1] & 
                                 myinput$published_year <= input$sliderYear[2]),]
  get_group_onlyDD = get_group_DD[which(get_group_DD$category == "DD"),]
  group_rows_DD = nrow(get_group_DD)
  group_rows_onlyDD = nrow(get_group_onlyDD)
  proportion = round((group_rows_onlyDD/group_rows_DD ) *100,digits = 1)
  
  myicon = ifelse(proportion > 40, "fa-arrow-up","arrows-alt-h")
  mycolour = ifelse(proportion > 40, "#d5392a", "#acacac")
  val = proportion
  
  valueBox(val, 
           #icon = myicon, 
           color =  mycolour)
})
```

### Increase in number of assessments since `r max(all_plants_sp$published_year)-1 `

```{r}
renderValueBox({
  get_group = get(input$Group)
  
  group_rows = nrow(subset(get_group, subset = published_year == max(published_year)))
  
  valueBox(prettyNum(group_rows, big.mark = ","), icon = "fa-arrow-up", color = "#acacac")
})
```

### Number of assessments > 10 years old

```{r}
renderValueBox({
  get_group = get(input$Group)
  
  group_rows = nrow(subset(get_group, subset = published_year < max(published_year)-10))
  
  valueBox(prettyNum(group_rows, big.mark = ","), color = "#acacac")
})
```

Trends {data-orientation=rows}
=====================================     
  
  Inputs{.sidebar}
-------------------------------------
  
  Red List Version 2018-1

- [IUCN Red List](http://www.iucnredlist.org/)




**Left tab** shows count of assessments per year

**Middle tab** shows breakdown of assessments categorised as 'new' (assessed within 5 years of being described) and 'old' (assessed > 5 years after description)

**Right tab** shows trends in use of different assessment languages






Row {data-height=600 .tabset .tabset-fade}
-------------------------------------
### Assessment count per year
  
  ```{r, }

### Count of Red List Assessments for plant species over time (Kingdom = PLANTAE)
all_plants_sp_pubyear = subset(all_plants_sp, published_year >2002 & published_year < 2018)

# filter out <2003 (old assessments, old criteria) and >2017 (2018 not finished yet)

ggplotly(
  ggplot(all_plants_sp_pubyear, aes(as.factor(published_year)))+ 
    geom_bar(fill = "#FF6666") +
    theme_minimal() +
    labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "")
)
```

### Species description vs Red List assessment date

```{r, }
diffplot <- read_csv("diffplot.csv", guess_max = 50000)

# adding in a single record to get a point for 2005 - must be another way to do this...
#extralev = data.frame(published_year = "2005", age = "Unknown")
#diffplot = rbind(diffplot, extralev)

#tax.nov.plot = subset(tax.nov.year, tax.nov.year$Year >="2003" & tax.nov.year$Year < "2018")

# now summarise the ratio of old and new for each red list assessment year
library(ggplot2)
fig_2 = ggplot() + 
  geom_bar(data = diffplot, aes(x = as.factor(published_year), fill = as.factor(age)), 
           position = position_dodge(preserve = 'single')) + 
  
  labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "") +
  theme_minimal() +
  scale_fill_manual(values = c("New" = "#f37735", "Old" = "#ffc425", "Unknown" = "light grey")) +
  #geom_point(data = tax.nov.plot, aes(x = as.factor(Year), y = Count), stat = "identity", size = 2, shape = 16) +
  #scale_colour_manual(name = "Year", labels = "test", values = "black") +
  scale_x_discrete(expand = c(0, 0)) + scale_y_continuous(expand = c(0,100)) 

ggplotly(
  fig_2
)
```


### Assessment language
```{r, }

# show bar with breakdown by language (fill)
# get language from new red list download page - import file
# ask ackbar to add this to API?

# import the Red List download
assessments_plants_2018_1 <- read_csv("assessments_plants_2018_1.csv", guess_max = 50000)

# assign the NAs as English
assessments_plants_2018_1$language[is.na(assessments_plants_2018_1$language)] <- "English"

# query out pre 2003
assessments_plants_2018_1 = subset(assessments_plants_2018_1, yearPublished >2002)

# plot
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 
rl_language = ggplot(data = assessments_plants_2018_1, aes(x = factor(yearPublished), fill = language))
rl_language = rl_language + geom_bar() +
  theme_minimal() +
  scale_fill_manual(values=cbPalette) +
  labs(x = "Red List Publication Year", y = "Number of plant species assessments", fill = "")

ggplotly(
  rl_language
)

```

